---
title: "Mid-term project instructions"
author: "Austin Winkler"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
---


```{r}
# Load necessary libraries
install.packages("lme4")
library(lme4)
```

```{r}
# Read data
IHdata <- read.csv("../Data/Iron_Horse_Stats_2.csv")

IHdata
```

```{r}
# One-way ANOVA
model <- aov(yield_buac ~ Nrate_lbac, data = data)
summary(model)

# Check assumptions with plots
par(mfrow = c(2, 2))
plot(model)

```
Residuals vs. Fitted: To see if there are any obvious patterns (indicating non-linearity) or non-constant variance of residuals (indicating heteroscedasticity).
Normal Q-Q: To check if the residuals are approximately normally distributed.
Scale-Location (also called Spread-Location): To check for homoscedasticity.
Residuals vs. Leverage: To check for influential outliers.


Quantile Regression
```{r}
install.packages("quantreg")
library(quantreg)

# Fit a quantile regression model for the median (0.5 quantile)
qr_model <- rq(yield_buac ~ Nrate_lbac, data = data, tau = 0.5)
summary(qr_model)

```



```{r}
Anova(yield_model, type=3)
```






