---
title: "Mid-term project instructions"
author: "Austin Winkler & Taranjeet Sodhi"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
---

# Mid-term project GitHub repository URL:  

Paste here the URL of the GitHub project (after going through the instruction steps)

# Instructions  
This file contains both the **instructions** for the mid-term project and placeholders for your code. You are required to use this file to produce code, output, and answers to the questions below.  

Besides simply creating output, make sure to interpret the output. You will need to create tables and/or plots to arrive at the answers, and then comment on what you found below it.    

To get you setup, you will need to:  

  - Student #1: create a repository on your GitHub account. You can call this repository "2024dsa_midproject_groupX", **where X is the number of your group**. Make it public, add a README, add a .gitignore for R.  
  - Student #1: follow the steps we did in class to start a new RStudio project with version control.  
  - Student #1: in your computer, create the sub-folders code, data, output, and move your data set into the `data` folder. Also, student 1 moves this current script into the `code` folder. Do a git stage, commit, push.  
  - Student #1: on GitHub, go the repository settings and invite your partner to be a collaborator in the repository. That will give them push permission.  
  - Now, both students should clone this repository on their computers like we did in class. Make sure this step works well and that you can pull and push from GitHub.  
  - Student 2, after cloning, does a git pull to get all these updates on their computer.  
  - Student 1 and 2 work together to update the README file. README files should explain what the repository is about, the goals of that project, who is working in it, and any other important details you may find.  
  
# Introduction  
Describe here the introduction of your problem. Consider this as a shortened version of your paper, where you will briefly discuss in 3-4 paragraphs what is the issue/gap in literature, and how the data you collected will help answer this gap.  

Efficient nitrogen (N) fertilizer use is crucial for the economic and environmental sustainability of corn production in Georgia. The challenge is achieving optimal corn growth without the economic losses associated with over-application, which also threatens environmental health through nitrate leaching, or under-application, which risks reduced yields. Current reliance on the Crop Yield Goal (CYG) method, with its uniform application recommendations across diverse agricultural contexts, fails to address the state's varying soil types, climates, and irrigation capabilities, highlighting a significant gap in the development of region-specific N recommendations.

The literature underscores a lack of tailored N fertilizer guidance that considers the unique agricultural landscapes across Georgia. The CYG method's shortcomings, including its reliance on historical yields and disregard for economic variables like fluctuating N fertilizer and grain prices, limit its effectiveness. This points to a pressing need for an approach that better aligns N application rates with the dynamic factors influencing crop production.

Our research seeks to fill this gap by evaluating the current N recommendation methods against Georgia's agricultural diversity. Through corn N rate trials at ten locations statewide, we aim to create a dataset that reflects the relationship between N rates and corn yield under varied conditions. This data will underpin the development of the Georgia N Rate Calculators, tools intended to offer economically and environmentally sound, region-specific N recommendations. This approach promises to reconcile economic efficiency with environmental stewardship, addressing both the literature gap and the practical needs of Georgia's corn growers.

# Hypothesis and objectives  
Describe here your hypothesis, followed by your objectives. Make sure your hypothesis are testable and bold, and objectives are clear.  

We hypothesize that corn yield responses to nitrogen application rates above 201 kg N/ha will not be significantly different, suggesting that the Agronomic Optimum Nitrogen Rate (AONR) lies at or below this threshold. 
This study aims to demonstrate that a nitrogen recommendation model integrating regional specificity, economic considerations, and management practices will provide more precise and sustainable nitrogen rates without a statistically significant yield benefit from higher applications.

# Material and Methods  
Describe here your overall material and methods as it pertains to the analysis you will conduct, including study description, site/setup description, what equipment was used, etc. just like you would in a paper. Make sure to clearly explain what was measured and how.

In our study on Corn Nitrogen Recommendation Methods in Georgia, we evaluated nitrogen (N) fertilizer application strategies by comparing the Crop Yield Goal (CYG) method with the Agronomic Optimum Nitrogen Rate (AONR) and Economic Optimum Nitrogen Rate (EONR) across twelve diverse locations within the state. The research was designed to assess eight distinct N application rates (ranging from 0 to 470 K N/ha), replicated three times at each site to determine the most effective N rate for corn production across various Georgian agricultural environments. This methodical approach, encompassing locations like Tifton, Midville, Plains, Camilla, and Watkinsville, aims to tailor N fertilizer recommendations to the state's unique regional characteristics, thereby addressing the demand for precision in agricultural practices. For this project we are only considering midville. 

The methodology employed a randomized complete block design to ensure unbiased field data, with a mix of conventional and strip till practices reflecting the state's common farming techniques. Planting configurations and nitrogen applications were standardized across all test plots, except those designated for zero N, to facilitate an accurate assessment of Nitrogen Use Efficiency. Comprehensive data collection included weather monitoring, plant sampling at the R6 stage for nitrogen content analysis, and yield component measurements, with a special focus on irrigation's role in optimizing N use. The culmination of this rigorous data analysis seeks to refine N application recommendations for Georgia's corn production, enhancing both the economic return for farmers and the environmental sustainability of their practices.

## Study design  
Clearly describe your study design here, including treatment design (which factors and levels, the hierarchy among them, etc.), and your experimental design (number of reps/blocks, how was randomization performed, etc.), as we talked about in class.  

Our study design integrates comprehensive evaluations of nitrogen (N) fertilizer application methods to optimize corn production in Georgia, comparing the traditional Crop Yield Goal (CYG) method against the Agronomic Optimum Nitrogen Rate (AONR) and Economic Optimum Nitrogen Rate (EONR). The primary aim is to refine N fertilizer recommendations through a meticulous examination of different N rates' effects on corn yield, accounting for regional variability across the state's diverse agricultural landscapes.

Treatment Design: The core of our treatment design involves the application of eight distinct N rates: 0, 67, 134, 201, 269, 336, 403, 470 K N/ac. These rates were chosen to span a broad spectrum from no nitrogen application to a high application rate, enabling us to assess the impact of varying N levels on corn yield and Nitrogen Use Efficiency. This range allows for the analysis of both agronomic and economic optimization of N use, aligning with the objectives of comparing AONR and EONR methods against CYG recommendations. The treatments were structured to capture the effects of N application across different environmental conditions and management practices inherent to Georgia's agricultural regions.

Experimental Design: Our study meticulously examined nitrogen (N) fertilizer application's impact on corn yield across eight diverse locations in Georgia, employing a randomized complete block design to ensure comprehensive and unbiased analysis. By applying eight distinct N rates (0 to 470 K N/ac) replicated three times at each site, we aimed to span the entire spectrum of possible N applications, from none to high. This structure allowed us to rigorously test the Agronomic Optimum Nitrogen Rate (AONR) and Economic Optimum Nitrogen Rate (EONR) methods against the conventional Crop Yield Goal (CYG) recommendations. Randomization within each block was critical to minimizing location-specific environmental effects and ensuring that each N rate treatment had an equal probability of assignment across the plots, thereby enhancing the reliability of our findings on the optimal N usage for maximizing corn production efficiency and profitability in Georgia's varied agricultural contexts.

## Statistical analysis  
Describe here your statistical analysis, including what type of ANOVA model you ran (based on the design above), what was your response variable, what were your explanatory variables and how were the explanatory variables treated (random or fixed). Provide your alpha level. Explain which function from which package you used to analyze this data. Explain how you checked linear model assumptions and whether or not they were met. Overall, make sure you explain in sufficient detail that, if given your data, a knowledgeable person would be able to reproduce your analysis exactly. 

The ANOVA model that was ran was the type 3 ANOVA model. The response variable yield. The explanatory variables was the different Nitrogen rates. The explanatory variables were treated with a randomized complete block design. The alpha level was 0.05. The LMER function was used from the LME4 package. The following condition were checked: residual independence, homoscedascity, redidual nomadity, and residual outliers. All of those conditions were met.


# Results  
Here is where the coding is going to happen, and it will be completely up to you. Include under this section as many sub-sections (using ##) and as many chunks needed to create the analytical workflow for your analysis, starting at loading packages and data, wrangling, EDA, modeling, assumptions checking, ANOVA table, means, pairwise comparisons, and final publication-quality plot.  

##Loading packages
```{r setup}
#install.packages("tidyverse")
#install.packages("tidyr")
#install.packages("dplyr")
#install.packages("ggplot2")
#install.packages("openxlsx")
#install.packages("car")
#install.packages("broom")
#install.packages("emmeans")
#install.packages("multcomp")
#install.packages("rlang")
library(tidyr)
library(dplyr)
library(ggplot2)
library(janitor)
library("car")
library("broom")
library(emmeans)
library(multcomp)
library(tidyverse) # for data wrangling and plotting
library(openxlsx)
```

##Loading data
```{r}
data <- read.csv("../Data/Midville_Stats.csv")

data

```
N fertilizer rates: 0, 67, 134, 201, 269, 336, 403, 470

## Data Wrangling
```{r}
data_w <- data %>%
  clean_names()%>%
  mutate(trtnum=1:nrow(.)) %>%
  slice(-49) %>%
  mutate(rep = factor(rep(1:6, times = 8)))%>%
  dplyr::select(plot, rep, nrate_kgha, yield_kgha)%>%
  mutate(nrate_kgha = factor(nrate_kgha))

data_w
```
## Data summary
```{r}
summary(data_w)
```
## Plotting data
```{r}
ggplot(data_w, aes(x = nrate_kgha, 
                    y = yield_kgha,
                    color = nrate_kgha)) +
  geom_boxplot() +
  geom_jitter() +
  theme(legend.position = "none")



```
## Model fitting and summary
```{r model}
# Changing to sum-to-zero contrast
options(contrasts = c("contr.sum", "contr.poly"))

# Model fitting
yield_model <-lm(yield_kgha~ rep + nrate_kgha,
              data=data_w)

yield_model
# Summary
summary(yield_model)
```
## Running Anova
```{r ANOVA}
Anova(yield_model, type=3)
```
# Linear model assumptions  
## Extracting residuals
First, let's extract our model residuals, and also create studentized residuals.  
```{r rcbd_resid}
rcbd_resid <- augment(yield_model) %>%
  mutate(.studresid=rstudent(yield_model))

rcbd_resid
```
Now, linear model assumptions:  

- Residual independence (no pattern)  
- Residual variance homogeneity (homoscedasticity)  
- Residual normality (discuss this!)  
- Outlier detection


1. Independence 
## Checking Independence
```{r }
ggplot(rcbd_resid, aes(x=.fitted, y=.studresid))+
  geom_hline(yintercept = 0, color="red")+
  geom_point(shape = 21,
             fill = "purple", 
             size = 3,
             alpha = .7)+
  #geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "red")+
  theme_bw()
```
2. Residual Homoscedascity 

## Residual Homoscedascity 
```{r }
ggplot(rcbd_resid, aes(x=.fitted, y=.studresid))+
  geom_hline(yintercept = 0, color="red")+
  geom_point(shape = 21,
             fill = "purple", 
             size = 3,
             alpha = .7)+
  geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "red")+
  theme_bw()

```
3. Residual Normality

## Residual Normality
```{r}
ggplot(rcbd_resid, aes(sample=.studresid))+
  stat_qq(  shape = 21,
            fill = "purple", 
            size = 3,
            alpha = .7
  )+
  stat_qq_line()+
  labs(x = "Theoretical quantile",
       y = "Sample quantile")+
  theme_bw()
```

## Density Plot
```{r}
ggplot(rcbd_resid, aes(x=.studresid))+
  geom_density(color = "black",
               fill = "purple",
               alpha = .7)+
  scale_x_continuous(breaks = c(-3,0,3), limits = c(-3,3))+
  theme_bw()


```

## Residual outliers 
```{r}
ggplot(rcbd_resid, aes(x=.fitted, y=.studresid))+
  geom_hline(yintercept = 0, color="red")+
  geom_point(shape = 21,
             fill = "purple", 
             size = 3,
             alpha = .7)+
  geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "red")+
  theme_bw()

```
## model means 
```{r interaction means all}
rcbd_means_all <-emmeans(yield_model, ~nrate_kgha)

rcbd_means_all
```
## pairwise comparison 
```{r interaction pwc all}
rcbd_cld_all <- cld(rcbd_means_all, 
                   reversed=T, 
                   adjust="none",
               Letters=letters,
               alpha = 0.05)

rcbd_cld_all
```

## Wrangling
```{r selected pwc}
rcbd_cld_selected <- rcbd_cld_all%>%
  as.data.frame()%>%
  mutate(letter=trimws(.group))
  

rcbd_cld_selected
```
## Final plot
```{r rcbd final plot}

ggplot(mapping = aes(fill = nrate_kgha))+
  geom_boxplot(data=yield_model,
               aes(x=nrate_kgha,
                  y=yield_kgha),
               alpha=.8
                    )+
  geom_jitter(data=yield_model,
              aes(x= nrate_kgha,
                  y=yield_kgha
                  ),
              shape=21,
              alpha=.6,
              size=3
              )+
  geom_label(data=rcbd_cld_selected,
             aes(x= nrate_kgha,
                 y= emmean,
                 label=letter),
            fill="grey"
                 )+
  scale_fill_viridis_d()+
  theme(legend.position = "none",
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y=element_blank()
        
        )
  
```
This graph shows the variation in crop yield (kg/ha) against nitrogen application rates (kg/ha). There's a noticeable trend of increasing yield with rising nitrogen rates up to a threshold, after which yields plateau, indicating an optimal nitrogen application range. The box plots reveal a widening distribution of yields at higher nitrogen levels, highlighting the variability and potential risk of exceeding the optimal nitrogen application rate.


Make sure to run a model that reflects your study design. Even if your study design does not include one of the designs covered in class, you are still expected to run the most appropriate model. If you need help for references, let me know.  

Before each chunk, describe the steps being performed in that chunk. For example, "Here I will load the data".  

If a chunk produces output, like printing a data frame, statistical summary, a plot, ANOVA table, etc., make sure to write text interpreting what you see and how you can/will use that information to move forward to the next steps in the workflow.  

# Team work in GitHub  
Whether you are working with your future-self or as duos, make sure to stage, commit, and push after finishing each of the sub-sections above. When committing, write commit messages that are short and descriptive (e.g., finished wrangling).  

If you are working in duos, make sure to split the workload. I will check your collaborative work through the commit history, and if one student has contributed significantly more than the other, than that will impact grades.  

**Tip 1**: to avoid merge conflicts, make sure to **pull** first, and then start working locally. That will ensure that any changes made by your partner will be "downloaded" before you make changes to the files locally.  

**Tip 2**: make use of the Issues on this repository to set up to-do lists and assign tasks to different people. You can also use each issue/task to discuss how things should be run and get to an agreement.  

# Submitting your work  
Once you have developed all the code and answers, make sure to Render this quarto file.  

**Notes on rendering**:  

- Make sure to render your work and inspect how the final html look like.  
- If it does not look professional for whatever reason, then fix the issue, re-render it, recheck.  
- Only send me your work once your html file looks professional.  
- Some potential issues you may encounter and how to fix them:   
  - Some times your code may be creating a table output that is waaay to long and cumbersome to scroll through when rendered. If this is the case, make it more professional looking by using the `head()` function to only print the first handful of rows (instead of thousands of rows).  
  
  - **DO NOT** delete the file's heading levels (# and ##). They set up the proper heading 1 and 2 levels, and I use them to guide my grading.  
  
  - If a given chunk is also outputting warnings or messages, inhibit this behavior by changing the chunk options `message` and `warning` to `FALSE`.  
  
  - If, after rendered, 2 lines of text are connected and you wish to "break line" between them, add 2 extra spaces after the first one.  

After rendering, an .html file will be created on your `code` folder.  

Rename this file to `LASTNAME1-LASTNAME2_midtermproject.html`.    
For ex., `Bastos-Mendes_midtermproject.html`.

Send the html file to my email (lmbastos@uga.edu) by **April 11th** 11:59 pm.  













  

  

