---
title: "Mid-term project instructions"
author: "Taranjeet Sodhi"
format:
  html:
    code-fold: false
    embed-resources: true
    toc: true
    number-sections: true
    theme: cerulean
    
    
    
    
---
```{r setup}
# Loading packages
install.packages("tidyverse")
install.packages("tidyr")
install.packages("dplyr")
install.packages("ggplot2")
install.packages("openxlsx")
install.packages("car")
install.packages("broom")
install.packages("emmeans")
library(tidyr)
library(dplyr)
library(ggplot2)
library(janitor)
library("car")
library("broom")
library(emmeans)


library(tidyverse) # for data wrangling and plotting
library(openxlsx)
```
```{r}
data <- read.csv("../Data/Midville_Stats.csv")

data

```
N fertilizer rates: 0,60,120,180,240,300,360,420

# Wrangling
```{r}
data_w <- data %>%
  clean_names()%>%
  dplyr::select(plot, nrate_kgha, yield_kgha)%>%
  mutate(trtnum=1:nrow(.))

data_w
```
# summary
```{r}
summary(data_w)
```

```{r}
ggplot(data_w, aes(x = nrate_kgha, 
                    y = yield_kgha,
                    color = nrate_kgha)) +
  geom_boxplot() +
  geom_jitter() +
  theme(legend.position = "none")



```

```{r model}
# Changing to sum-to-zero contrast
options(contrasts = c("contr.sum", "contr.poly"))

# Model fitting
yield_model <-lm(yield_kgha~nrate_kgha,
              data=data_w)

yield_model
# Summary
summary(yield_model)
```

```{r ANOVA}
Anova(yield_model, type=3)
```
# Linear model assumptions  
## Extracting residuals
First, let's extract our model residuals, and also create studentized residuals.  

```{r rcbd_resid}
rcbd_resid <- augment(yield_model) %>%
  mutate(.studresid=rstudent(yield_model))

rcbd_resid
```
Now, linear model assumptions:  

- Residual independence (no pattern)  
- Residual variance homogeneity (homoscedasticity)  
- Residual normality (discuss this!)  
- Outlier detection


1. Independence 
```{r }
ggplot(rcbd_resid, aes(x=.fitted, y=.studresid))+
  geom_hline(yintercept = 0, color="red")+
  geom_point(shape = 21,
             fill = "purple", 
             size = 3,
             alpha = .7)+
  #geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "red")+
  theme_bw()
```
2. Residual Homoscedascity 


```{r }
ggplot(rcbd_resid, aes(x=.fitted, y=.studresid))+
  geom_hline(yintercept = 0, color="red")+
  geom_point(shape = 21,
             fill = "purple", 
             size = 3,
             alpha = .7)+
  geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "red")+
  theme_bw()

```
3. Residual Normality

```{r}
ggplot(rcbd_resid, aes(sample=.studresid))+
  stat_qq(  shape = 21,
            fill = "purple", 
            size = 3,
            alpha = .7
  )+
  stat_qq_line()+
  labs(x = "Theoretical quantile",
       y = "Sample quantile")+
  theme_bw()
```


```{r}
ggplot(rcbd_resid, aes(x=.studresid))+
  geom_density(color = "black",
               fill = "purple",
               alpha = .7)+
  scale_x_continuous(breaks = c(-3,0,3), limits = c(-3,3))+
  theme_bw()


```

# Residual outliers 
```{r}
ggplot(rcbd_resid, aes(x=.fitted, y=.studresid))+
  geom_hline(yintercept = 0, color="red")+
  geom_point(shape = 21,
             fill = "purple", 
             size = 3,
             alpha = .7)+
  geom_smooth()+
  geom_hline(yintercept = c(-3,3), color = "red")+
  theme_bw()

```
# model means 
```{r interaction means all}
rcbd_means_all <-emmeans(rcbd_resid, ~nrate_kgha:krate_kgha)

rcbd_means_all
```

